Bird beak's elaboration and exploration

### 2021-07-14 Analysis/07-benchmarking.Rmd
Update 07-benchmarking.Rmd

Saving parameters:
 * `pr = TRUE/FALSE`: save the posteriors of the random effects (FALSE?)
 * `pl = TRUE/FALSE`: save the posteriors of the latent variable (TRUE?)
 * `DIC = TRUE/FALSE`: deviance and deviance information criterion are calculated
 * `saveX = TRUE/FALSE`: save fixed effect design matrix
 * `saveZ = TRUE/FALSE`: save random effect design matrix
 * `saveXL = TRUE/FALSE`: save structural parameter design matrix

@@

```{r, run param chains}
## Set up the parametrising chains models
param_10g_2d <- make.mini.chains(data         = chara_data,
                                   dimensions   = c(1:2),
                                   tree         = chara_tree,
                                   trait.family = "gaussian",
                                   randoms      = c("global", "clade"),
                                   residuals    = "global",
                                   priors       = 0.02,
                                   verbose      = TRUE,
                                   parameters   = list(
                                       nitt   = 10,
                                       thin   = 1,
                                       burnin = 0))
## Set up the parametrising chains models
param_20g_2d <- make.mini.chains(data         = chara_data,
                                   dimensions   = c(1:2),
                                   tree         = chara_tree,
                                   trait.family = "gaussian",
                                   randoms      = c("global", "clade"),
                                   residuals    = "global",
                                   priors       = 0.02,
                                   verbose      = TRUE,
                                   parameters   = list(
                                       nitt   = 20,
                                       thin   = 1,
                                       burnin = 0))
## Set up the parametrising chains models
param_30g_2d <- make.mini.chains(data         = chara_data,
                                   dimensions   = c(1:2),
                                   tree         = chara_tree,
                                   trait.family = "gaussian",
                                   randoms      = c("global", "clade"),
                                   residuals    = "global",
                                   priors       = 0.02,
                                   verbose      = TRUE,
                                   parameters   = list(
                                       nitt   = 30,
                                       thin   = 1,
                                       burnin = 0))
## Set up the parametrising chains models
param_40g_2d <- make.mini.chains(data         = chara_data,
                                   dimensions   = c(1:2),
                                   tree         = chara_tree,
                                   trait.family = "gaussian",
                                   randoms      = c("global", "clade"),
                                   residuals    = "global",
                                   priors       = 0.02,
                                   verbose      = TRUE,
                                   parameters   = list(
                                       nitt   = 40,
                                       thin   = 1,
                                       burnin = 0))
## Set up the parametrising chains models
param_50g_2d <- make.mini.chains(data         = chara_data,
                                   dimensions   = c(1:2),
                                   tree         = chara_tree,
                                   trait.family = "gaussian",
                                   randoms      = c("global", "clade"),
                                   residuals    = "global",
                                   priors       = 0.02,
                                   verbose      = TRUE,
                                   parameters   = list(
                                       nitt   = 50,
                                       thin   = 1,
                                       burnin = 0))
param_list <- list(param_10g_2d,
                   param_20g_2d,
                   param_30g_2d,
                   param_40g_2d,
                   param_50g_2d)
```

```{r }
## Recording the results
system(paste("echo \"thin, iterations, dimensions, time, d_time, r_time, size, d_size, r_size\" > recording.csv"))

@@

## Saving the results
parameters <- paste(1, 10, 2, sep = ", ")
data <- paste(timer_base, 0, timer_base/10, chain_size_base, 0, chain_size_base/10, sep = ", ")
system(paste("echo \"", paste(parameters, data, sep = ", "), "\" >> recording.csv"))

## Other iterations
## Difference in iterations
d_itt <- 10
timer <- system.time(param_chains <- run.mini.chains(param_list[[2]], replicates = 1))["elapsed"]
chain_size <- object.size(param_chains)
parameters <- paste(1, 10 + d_itt, 2, sep = ", ")
data <- paste(timer, timer-timer_base, (timer-timer_base)/d_itt, chain_size, chain_size-chain_size_base, (chain_size-chain_size_base)/d_itt, sep = ", ")
system(paste("echo \"", paste(parameters, data, sep = ", "), "\" >> recording.csv"))

```

>>>>

> Here I'm not going to test the saving parameters since it's pretty much a constant RAM saving. By removing all the un-necessary stuff to save, we have 4.2 times smaller output!

@@

```{r}
## Function for setting up the parameters
set.params <- function(dimensions, thin, nitt) {
  ## Set up the parametrising chains models
  param1 <- make.mini.chains(data         = chara_data,
                             dimensions   = c(1:dimensions),
                             tree         = chara_tree,
                             trait.family = "gaussian",
                             randoms      = c("global", "clade"),
                             residuals    = "global",
                             priors       = 0.02,
                             verbose      = TRUE,
                             parameters   = list(
                                 nitt   = nitt*1,
                                 thin   = thin,
                                 burnin = 0))
  ## Set up the parametrising chains models
  param2 <- make.mini.chains(data         = chara_data,
                             dimensions   = c(1:dimensions),
                             tree         = chara_tree,
                             trait.family = "gaussian",
                             randoms      = c("global", "clade"),
                             residuals    = "global",
                             priors       = 0.02,
                             verbose      = TRUE,
                             parameters   = list(
                                 nitt   = nitt*2,
                                 thin   = thin,
                                 burnin = 0))
  ## Set up the parametrising chains models
  param3 <- make.mini.chains(data         = chara_data,
                             dimensions   = c(1:dimensions),
                             tree         = chara_tree,
                             trait.family = "gaussian",
                             randoms      = c("global", "clade"),
                             residuals    = "global",
                             priors       = 0.02,
                             verbose      = TRUE,
                             parameters   = list(
                                 nitt   = nitt*3,
                                 thin   = thin,
                                 burnin = 0))
  ## Set up the parametrising chains models
  param4 <- make.mini.chains(data         = chara_data,
                             dimensions   = c(1:dimensions),
                             tree         = chara_tree,
                             trait.family = "gaussian",
                             randoms      = c("global", "clade"),
                             residuals    = "global",
                             priors       = 0.02,
                             verbose      = TRUE,
                             parameters   = list(
                                 nitt   = nitt*4,
                                 thin   = thin,
                                 burnin = 0))
  ## Set up the parametrising chains models
  param5 <- make.mini.chains(data         = chara_data,
                             dimensions   = c(1:dimensions),
                             tree         = chara_tree,
                             trait.family = "gaussian",
                             randoms      = c("global", "clade"),
                             residuals    = "global",
                             priors       = 0.02,
                             verbose      = TRUE,
                             parameters   = list(
                                 nitt   = nitt*5,
                                 thin   = thin,
                                 burnin = 0))
  return(list(param1,
              param2,
              param3,
              param4,
              param5))
}
## Function for running the test
test.params <- function(dimensions, thin, nitt) {
  ## Setting up the parameters
  param_list <- set.params(dimensions = dimensions, thin = thin, nitt = nitt)

@@

  ## Preping the results
  parameters <- paste(thin, nitt, dimensions, sep = ", ")
  data <- paste(timer_base, 0, timer_base/nitt, chain_size_base, 0, chain_size_base/nitt, sep = ", ")
  ## Appending the file
  system(paste("echo \"", paste(parameters, data, sep = ", "), "\" >> recording.csv"))

  ## Second run
  timer <- system.time(param_chains <- run.mini.chains(param_list[[2]], replicates = 1))["elapsed"]
  chain_size <- object.size(param_chains)

  ## Preping the results
  parameters <- paste(thin, nitt*2, dimensions, sep = ", ")
  data <- paste(timer, timer-timer_base, (timer-timer_base)/nitt*1, chain_size, chain_size-chain_size_base, (chain_size-chain_size_base)/nitt*1, sep = ", ")
  ## Appending the file
  system(paste("echo \"", paste(parameters, data, sep = ", "), "\" >> recording.csv"))

  ## Third run
  timer <- system.time(param_chains <- run.mini.chains(param_list[[3]], replicates = 1))["elapsed"]
  chain_size <- object.size(param_chains)

  ## Preping the results
  parameters <- paste(thin, nitt*3, dimensions, sep = ", ")
  data <- paste(timer, timer-timer_base, (timer-timer_base)/nitt*2, chain_size, chain_size-chain_size_base, (chain_size-chain_size_base)/nitt*2, sep = ", ")
  ## Appending the file
  system(paste("echo \"", paste(parameters, data, sep = ", "), "\" >> recording.csv"))

  ## Fourth run
  timer <- system.time(param_chains <- run.mini.chains(param_list[[4]], replicates = 1))["elapsed"]
  chain_size <- object.size(param_chains)

  ## Preping the results
  parameters <- paste(thin, nitt*4, dimensions, sep = ", ")
  data <- paste(timer, timer-timer_base, (timer-timer_base)/nitt*3, chain_size, chain_size-chain_size_base, (chain_size-chain_size_base)/nitt*3, sep = ", ")
  ## Appending the file
  system(paste("echo \"", paste(parameters, data, sep = ", "), "\" >> recording.csv"))

  ## Fifth run
  timer <- system.time(param_chains <- run.mini.chains(param_list[[5]], replicates = 1))["elapsed"]
  chain_size <- object.size(param_chains)

  ## Preping the results
  parameters <- paste(thin, nitt*5, dimensions, sep = ", ")
  data <- paste(timer, timer-timer_base, (timer-timer_base)/nitt*4, chain_size, chain_size-chain_size_base, (chain_size-chain_size_base)/nitt*4, sep = ", ")
  ## Appending the file
  system(paste("echo \"", paste(parameters, data, sep = ", "), "\" >> recording.csv"))

  return(NULL)
}
```

## Running the tests for 1k generations (sampling every 100)

```{r, eval = TRUE}
## Initialising the results result table
system(paste("echo \"thin, iterations, dimensions, time, d_time, r_time, size, d_size, r_size\" > recording.csv"))

## Running the test
for(dim in 2:17) {
  test.params(dimensions = dim, thin = 100, nitt = 1000)
}
```

```{r}
## Upload the table
rable <- read.csv("recording.csv")

## Log values
run_time <- log(rable[, "time"])
run_size <- log(rable[, "size"])
cols <- rainbow(16)

par(mfrow = c(2,1))
plot(NULL, xlim = c(1, 5), ylim = range(run_time), main = "Run time", xlab = "10s of samples (per 1k iterations)", ylab = "log elapsed seconds")
for(i in 1:(nrow(rable)/5)) {
  lines(run_time[(1+5*(i-1)):(5+5*(i-1))], col = cols[i])
}

plot(NULL, xlim = c(1, 5), ylim = range(run_size), main = "Object size", xlab = "10s of samples (per 1k iterations)", ylab = "log bytes")
for(i in 1:(nrow(rable)/5)) {
  lines(run_size[(1+5*(i-1)):(5+5*(i-1))], col = cols[i])
}
```

## Running the tests for 10k generations (sampling every 1000)

```{r, eval = TRUE}
## Initialising the results result table
system(paste("echo \"thin, iterations, dimensions, time, d_time, r_time, size, d_size, r_size\" > recording2.csv"))

## Running the test
for(dim in 2:17) {
  test.params(dimensions = dim, thin = 1000, nitt = 10000)
}
```

```{r}
## Upload the table
rable <- read.csv("recording2.csv")

## Log values
run_time <- log(rable[, "time"])
run_size <- log(rable[, "size"])
cols <- rainbow(16)

par(mfrow = c(2,1))
plot(NULL, xlim = c(1, 5), ylim = range(run_time), main = "Run time", xlab = "10s of samples (per 10k iterations)", ylab = "log elapsed seconds")
for(i in 1:(nrow(rable)/5)) {
  lines(run_time[(1+5*(i-1)):(5+5*(i-1))], col = cols[i])
}

plot(NULL, xlim = c(1, 5), ylim = range(run_size), main = "Object size", xlab = "10s of samples (per 10k iterations)", ylab = "log bytes")
for(i in 1:(nrow(rable)/5)) {
  lines(run_size[(1+5*(i-1)):(5+5*(i-1))], col = cols[i])
}
```
